---
title: "Exercise 7"
editor: source
editor_options: 
  chunk_output_type: console
---

# Enrichment analysis

## Load Libraries
```{r}
#| warning: false
#| message: false
library(rGREAT)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
library(simplifyEnrichment)
library(flexclust)
```

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE)
```

## OverlapMatrix
```{r}
overlap <- readRDS("output/overlap_anno2.rds")
```

We have several cluster of genomic activities. We will run enrichment analysis of regions from one of them, "Active Promoters".

## Running enrichment analysis on regions of our interest

### subset the `GRanges`

```{r}
gr <- overlap[overlap$ATAC_RNA == "Active Promoters"]
seqlevels(gr) <- as.character(unique(seqnames(gr)))
```


### Running enrichment analysis
```{r}
res <- great(gr = gr, gene_sets = "BP", tss_source = "mm10")
```

## Results

### Volcano plot
```{r}
plotVolcano(res)
```


### Regions-gene assiciation plot
```{r}
plotRegionGeneAssociations(res)
```

### There is a `shiny` application to explore the results
```{r, eval=FALSE}
shinyReport(res)
```

### Simplyfied representation of GO terms

```{r}
tb <- getEnrichmentTable(res)
sig_go_ids <- tb$id[tb$p_adjust < 0.05]
cl <- simplifyGO(mat = sig_go_ids)
```


```{r}
summarizeGO(go_id = sig_go_ids, -log10(tb$p_adjust[tb$p_adjust < 0.05]), 
            axis_label = "average -log10(p.adjust)")
```

### Output
:::{.callout-important}
- The heatmap shows clear clustering of GO terms into two main clusters, indicating distinct functional groups.

- The GO terms on the right highlight biological processes relevant to neuronal and ion transport functions, such as transport, inorganic cation ion, exocytosis, synaptic, neurotransmitter, and behavior learning.

- The cluster with terms like cytosolic concentration, calcium ion suggests a calcium-related signaling or ion homeostasis process at play.

- The color gradient and similarity scores (up to ~0.6) indicate moderate similarity within clusters, which is reasonable for GO term groupings.
:::

### Interpretation
:::{.callout-important}
- E11.5 to E15.5 is a key window in neural development involving neuronal differentiation, synapse formation, and active signaling.

- Active promoters linked to ion transport, synaptic transmission, calcium signaling are expected to be regulated during this developmental stage.

- The clustering separating signaling and transport-related terms from behavior and learning terms also aligns with early neural circuit formation and functional maturation.
:::


## Question

**Perform enrichment analysis of any other category.**